---
title: "BIST8130_hw2"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
```

# Question 1
## Part b

What is the probability that at least 40 of these individuals will have at least one dental checkup?

```{r}
#assigning variables for number of trials, probability of success, and probability of failure
n = 56
p = 0.73
q_1 = 0.27
```

```{r}
#probability of x less than 40, P(X < 40)
P_X_less_40 = pbinom(39, n, p)
```

```{r}
#probability of x greater than or equal to 40, P(X >= 40)
P_X_ge_40 = 1 - P_X_less_40
```

The probability that at least 40 of these individuals will have at least one dental checkup is `r P_X_ge_40`.



## Part d

How many individuals do you expect to have at least one dental checkup?

```{r}
#calculating the number of individuals expected to have at least one dental checkup
expected_number = n * p
expected_rounded = floor(expected_number)
```

The expected number of individuals to have at least one dental checkup is `r expected_number` or `r expected_rounded` individuals.

## Part e

What is the standard deviation of the number of individuals who will have at least one dental checkup?

```{r}
#calculating the standard deviation of individuals who will have at least one dental checkup
sd_e = sqrt(n * p * q_1)
```

The standard deviation of the number of individuals who will have at least one dental checkup is `r sd_e`.


# Question 2

## Part a

What is the probability of having fewer than 3 tornadoes in the United States next year?

```{r}
#parameters for average number of tornadoes and threshold for fewer than 3 tornadoes
lambda = 6
q_2a = 2
```

```{r}
#calculating the probability of having fewer than 3 tornadoes in the United States next year
P_less_3 = ppois(q_2a, lambda)
```

## Part b

What is the probability of having exactly 3 tornadoes in the United States next year?

```{r}
#number of tornadoes
x = 3
```

```{r}
#calculating the probability of having exactly 3 tornadoes in the United States next year
P_exact_3 = dpois(x, lambda)
```

The probability of having exactly 3 tornadoes in the United States next year is `r P_exact_3`.

## Part c

What is the probability of having more than 3 tornadoes in the United States?

```{r}
#threshold for more than 3 tornadoes
q_2c = 3
```

```{r}
#calculating the probability of having more than 3 tornadoes in the United States
P_greater_3 = 1 - ppois(q_2c, lambda)
```

The probability of having more than 3 tornadoes in the United States is `r P_greater_3`.
