---
title: "BIST8130_hw2"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
```

# Question 1
## Part b

What is the probability that at least 40 of these individuals will have at least one dental checkup?

```{r}
#assigning variables for number of trials, probability of success, and probability of failure
n = 56
p = 0.73
q_1 = 0.27
```

```{r}
#probability of x less than 40, P(X < 40)
P_X_less_40 = pbinom(39, n, p)
```

```{r}
#probability of x greater than or equal to 40, P(X >= 40)
P_X_ge_40 = 1 - P_X_less_40
```

The probability that at least 40 of these individuals will have at least one dental checkup is `r P_X_ge_40`.



## Part d

How many individuals do you expect to have at least one dental checkup?

```{r}
#calculating the number of individuals expected to have at least one dental checkup
expected_number = n * p
expected_rounded = floor(expected_number)
```

The expected number of individuals to have at least one dental checkup is `r expected_number` or `r expected_rounded` individuals.

## Part e

What is the standard deviation of the number of individuals who will have at least one dental checkup?

```{r}
#calculating the standard deviation of individuals who will have at least one dental checkup
sd_1e = sqrt(n * p * q_1)
```

The standard deviation of the number of individuals who will have at least one dental checkup is `r sd_1e`.


# Question 2

## Part a

What is the probability of having fewer than 3 tornadoes in the United States next year?

```{r}
#parameters for average number of tornadoes and threshold for fewer than 3 tornadoes
lambda = 6
q_2a = 2
```

```{r}
#calculating the probability of having fewer than 3 tornadoes in the United States next year
P_less_3 = ppois(q_2a, lambda)
```

## Part b

What is the probability of having exactly 3 tornadoes in the United States next year?

```{r}
#number of tornadoes
x = 3
```

```{r}
#calculating the probability of having exactly 3 tornadoes in the United States next year
P_exact_3 = dpois(x, lambda)
```

The probability of having exactly 3 tornadoes in the United States next year is `r P_exact_3`.

## Part c

What is the probability of having more than 3 tornadoes in the United States?

```{r}
#threshold for more than 3 tornadoes
q_2c = 3
```

```{r}
#calculating the probability of having more than 3 tornadoes in the United States
P_greater_3 = 1 - ppois(q_2c, lambda)
```

The probability of having more than 3 tornadoes in the United States is `r P_greater_3`.


# Question 3

## Part a

What is the probability that a randomly selected American male between 20 and 29 years old has a systolic blood pressure above 137.0?
```{r}
#parameters for population mean, population standard deviation, and systolic blood pressure threshold
mean_3 = 128.0
sd_3 = 10.2
sbp_3a = 137.0
```

```{r}
#calculating the probability of systolic blood pressure > 137.0
P_above_137 = 1 - pnorm(sbp_3a, mean_3, sd_3)
```

The probability that a randomly selected American male between 20 and 29 years old has a systolic blood pressure above 137.0 is `r P_above_137`.

## Part b

What is the probability that the sample mean for blood pressure of 50 males between 20 and 29 years old will be less than 125.0?

```{r}
#sample size and sample mean
n_3b = 50
smean_3b = 125.0
```

```{r}
#calculating standard error of the mean
se_3b = sd_3 / sqrt(n_3b)
```

```{r}
#calculating z-score
zscore_3b = (smean_3b - mean_3) / se_3b
```

```{r}
#calculating the probability of sample mean < 125.0
P_smean_less_125 = pnorm(zscore_3b)
```

The probability that the sample mean for blood pressure of 50 males between 20 and 29 years old will be less than 125.0 is `r P_smean_less_125`.

## Part c

What is the 90th percentile of the sampling distribution of the sample mean X for a sample size of 40?

```{r}
#sample size
n_3c = 40
```

```{r}
#calculating standard error of the mean
se_3c = sd_3 / sqrt(n_3c)
```

```{r}
#calculating the 90th percentile
percentile_90 = qnorm(0.90, mean_3, se_3c)
```

The 90th percentile of the sampling distribution of the sample size mean X for a sample size of 40 is `r percentile_90`.